\input{style/sqtbeamer.tex}

\title{Intro to Quant}
\subtitle{Workshop}
\institute{Sydney Quant Traders \\ $\times$ \\ UNSW Competitive Programming \& Mathematics Society}
\date{}

\begin{document}

\maketitle

\begin{frame}{Workshop Outline}
  \begin{itemize}
    \item introduction to trading
    \begin{itemize}
      \item trades, orders, order matching
      \item overview of the industry
    \end{itemize}
    \item roles in trading, research and engineering
    \begin{itemize}
      \item job description
      \item interview process
      \item interview problems
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{Motivation}
  Why the trading industry?
  \begin{itemize}
    \item Intellectually challenging
    \item High agency and ownership
    \item Compensation
    \item Benefit to society
  \end{itemize}
\end{frame}

\section{Trading \& Markets}

\begin{frame}{Trading Introduction}
  We will introduce the concepts of
  \begin{itemize}
    \item trading
    \item exchanges
    \item the orderbook
  \end{itemize}

  \vspace{0.3cm}

  \textcolor{sqt-question}{
    I thought that firms don't care about prior knowledge of finance?
  }

  \begin{itemize}
    \item this is only true during the interview stage for large HFT firms
    \item beginner understanding of orderbook dynamics is useful for
    \begin{itemize}
      \item mock trading games
      \item general understanding of the industry
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{Spot Trades}
  \begin{sqt:tip}
    Alice says to Bob:
    \begin{center}
      ``I want to to sell you 5 AAPL shares. Each share will be sold at \$30. 
      This transfer will occur by the end of the next business day (tomorrow)." 
    \end{center}
  \end{sqt:tip}

  This defines a \textbf{spot trade} (as opposed to a \textbf{future trade}).

  \begin{sqt:definition}
    A \textbf{spot trade} is an agreement to exchange assets for money, with 
    the actual transfer occuring almost immediately (within 1 or 2 business days).
    \begin{itemize}
      \item sides (Alice is selling, Bob is buying)
      \item quantity of the asset being traded ($5$)
      \item price per asset ($\$30$)
    \end{itemize}
  \end{sqt:definition}
\end{frame}

\begin{frame}{Trade Execution and Settlement}
  Once Bob \textbf{agrees} to this trade, the trade can be \textbf{executed}.

  \begin{sqt:definition}
    Upon trade \textbf{execution}, both parties are given \textbf{settlement obligations}.
  \end{sqt:definition}
  \begin{sqt:tip}
    For example, in Alice \& Bob's equity spot trade, the obligations are:
    \begin{itemize}
      \item Alice must give Bob $5$ shares of AAPL by the settlement deadline
      \item Bob must give Alice $5\times \$30 = \$150$ by the settlement deadline
    \end{itemize}
    Once the obligations are fulfilled, the trade is considered \textbf{settled}.
  \end{sqt:tip}

  Obligations on other asset classes and types of trades (e.g. futures and options) 
  are considerably more complex than this.
  Concepts related to \textbf{clearing} and \textbf{settlement} are 
  considered part of the \textbf{post-trade life-cycle}.
\end{frame}

% TODO: finish
\begin{frame}{Holdings and Position}
  At each point in time, each participant has a \textbf{holdings} and a \textbf{position} in each asset.
  \begin{sqt:definition}
    \textbf{Position} includes the effects of executed \textbf{trades} that have not been \textbf{settled} yet.
  \end{sqt:definition}

  \begin{center}
    \input{assets/holdings_vs_positions.tex}
  \end{center}
\end{frame}

\begin{frame}{Holdings and Position}
  \begin{sqt:result}
    Your \textbf{exposure} and \textbf{PnL} is determined by your \textbf{position}.
  \end{sqt:result}

  This is because the terms of the trade (price, quantity) are locked in immediately upon trade execution,
  so the transfer of assets and money can be thought of us occuring immediately.

  \begin{sqt:tip}
    For example, you agree to buy a share of AAPL at \$30.
    Immediately, you are ``exposed'' to the movement of AAPL stock price.
    If the market price goes down to \$10, you are still paying \$30 for it upon settlement,
    but might end up having to sell it away for \$10.
  \end{sqt:tip}

  As a result, for trading games and simulations, we can abstract away the entire post-trade life-cycle
  and just consider our asset and cash \textbf{position}, disregarding \textbf{holdings} entirely.
\end{frame}

\begin{frame}{What is an Exchange}
  \begin{sqt:definition}
    An \textbf{exchange} is a centralised venue for participants to\\
    find willing counterparties to make \textbf{trades} with.
  \end{sqt:definition}
  \begin{sqt:definition}
    If a trade does not occur on an exchange, then the trade is \textbf{over-the-counter} (OTC).
  \end{sqt:definition}
  \vspace{0.5cm}

  \begin{center}
    \begin{columns}
      \begin{column}[t]{0.3\textwidth}
        US exchanges
        \begin{itemize}
          \item (XNYS) NYSE
          \item (XNAS) Nasdaq
          \item (XCBO) CBOE
          \item (XCME) CME
        \end{itemize}
      \end{column}

      \begin{column}[t]{0.3\textwidth}
        APAC exchanges
        \begin{itemize}
          \item (XHKG) HKEX
          \item (XSHG) SSE
          \item (XSGE) SHFE
        \end{itemize}
      \end{column}
    \end{columns}
  \end{center}
\end{frame}

\begin{frame}{Limit Orders}
  Each participant maintains a set of \textbf{limit orders}, which the exchange uses 
  to automatically infer whether they would agree to any hypothetical trade or not.\\
  Here is what a limit order may look like:
  \begin{sqt:tip}
    Alice tells the exchange
    \begin{center}
      ``I am willing to sell a maximum of 5 AAPL shares.\\
      I am willing to sell each share at a price of \$30 or better.''
    \end{center}
  \end{sqt:tip}

  This is called \textbf{inserting} an order.

  \begin{center}
    \input{assets/limit_order_intro.tex}
  \end{center}
\end{frame}

\begin{frame}{Bids and Offers}
  By submitting the \textbf{sell order}, Alice authorizes the exchange to accept trades on her behalf, 
  until 5 AAPL shares have been \textbf{sold}. Once this happens, the order is considered \textbf{filled}.

  \begin{sqt:definition}
    \begin{itemize}
      \item A \textbf{bid} (buy order) expresses a willingness to buy
      \item An \textbf{offer/ask} (sell order) expresses a willingness to sell
    \end{itemize}
  \end{sqt:definition}

  \begin{center}
    \input{assets/bids_and_asks.tex}
  \end{center}
\end{frame}

\begin{frame}{Orderbook}
  \begin{sqt:definition}
    The exchange keeps track of each participant's \textbf{open orders} (unfilled),\\
    and collects them into an \textbf{orderbook}.
  \end{sqt:definition}

  \begin{center}
    \input{assets/orderbook.tex}
  \end{center}
\end{frame}

\begin{frame}{Limit Orders In Cross}

  \begin{sqt:definition}
    Two limit orders are \textbf{in cross} if
    \begin{itemize}
      \item One order is a \textbf{bid} and the other is an \textbf{ask}.
      \item There exists an \textbf{intersection} between the price ranges they are willing to trade at
    \end{itemize}
  \end{sqt:definition}
  \begin{center}
    \input{assets/in_cross.tex}
  \end{center}
\end{frame}

\section{Order Matching}

\begin{frame}{Order Matching Algorithm}
  We will demonstrate \textbf{continuous matching} with \textbf{price-time priority}.\\
  This is a \textbf{matching algorithm} that determines the trades made when orders are inserted.

  \begin{itemize}
    \item Orders are \textbf{received} and then \textbf{handled} by the exchange one at a time.
    \item The order currently being \textbf{handled} is the \textbf{aggressing order}.
    \item The \textbf{aggressing order} is matched against \textbf{open orders} it is \textbf{in cross} with.
    \begin{itemize}
      \item Choose the \textbf{open order} according to \textbf{price-time priority}.
      \item Execute the trade at the price level of the \textbf{open order}, for the maximum quantity possible.
    \end{itemize}
    \item This matching occurs until either:
    \begin{itemize}
      \item The aggressing order runs out of quantity:
      \begin{itemize}
        \item The order is considered \textbf{filled} and is discarded.
      \end{itemize}
      \item No more orders \textbf{in cross} remain:
      \begin{itemize}
        \item The order is \textbf{unfilled} or only \textbf{partially filled}.
        \item If the order was a \textbf{fill-and-kill} (FAK), then the order is discarded.
        \item If the order was a \textbf{good-till-cancel} (GTC), then it is inserted at its price level at the back of the queue.
      \end{itemize}
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{Small GTC Buy Order: Initial}
  \begin{center}
    \input{assets/small_gtc_frame_0.tex}
  \end{center}
\end{frame}

\begin{frame}{Small GTC Buy Order: Animation}
  \begin{center}
    \only<1>{\input{assets/small_gtc_frame_1.tex}}
    \only<2>{\input{assets/small_gtc_frame_2.tex}}
    \only<3>{\input{assets/small_gtc_frame_3.tex}}
    \only<4>{\input{assets/small_gtc_frame_4.tex}}
    \only<5>{\input{assets/small_gtc_frame_5.tex}}
  \end{center}
\end{frame}

\begin{frame}{Small GTC Buy Order: Result}
  \begin{center}
    \input{assets/small_gtc_frame_result.tex}
  \end{center}
\end{frame}

\begin{frame}{Large GTC Buy Order: Initial}
  \begin{center}
    \input{assets/large_gtc_frame_0.tex}
  \end{center}
\end{frame}

\begin{frame}{Large GTC Buy Order: Animation}
  \begin{center}
    \only<1>{\input{assets/large_gtc_frame_1.tex}}
    \only<2>{\input{assets/large_gtc_frame_2.tex}}
    \only<3>{\input{assets/large_gtc_frame_3.tex}}
    \only<4>{\input{assets/large_gtc_frame_4.tex}}
    \only<5>{\input{assets/large_gtc_frame_5.tex}}
    \only<6>{\input{assets/large_gtc_frame_6.tex}}
  \end{center}
\end{frame}

\begin{frame}{Large GTC Buy Order: Result}
  \begin{center}
    \input{assets/large_gtc_frame_result.tex}
  \end{center}
\end{frame}

\begin{frame}{Aggregated Orderbook}
  \begin{sqt:result}
    Usually, exchanges only reveal to the participants:
    \begin{itemize}
      \item the set of \textbf{their own} open orders
      \item the \textbf{aggregated} orderbook (all quantities on each level are aggregated)
    \end{itemize}
  \end{sqt:result}

  \begin{center}
    \input{assets/aggregated_orderbook.tex}
  \end{center}
\end{frame}

\begin{frame}{Liquidity and Market Impact}
  \begin{sqt:definition}
    An orderbook with tight spreads and large volume is considered \textbf{liquid}.
  \end{sqt:definition}

  In an \textbf{illiquid} market, orders can \textbf{eat through} the orderbook.\\
  \textbf{Market impact} is when an action (order insertion) has an effect on the market price.

  \begin{center}
    \input{assets/market_impact.tex}
  \end{center}
\end{frame}

\begin{frame}{Limit Orderbook Examples}
  TODO: ditch this table, use the mathcha drawings instead for consistent style

  \begin{columns}[T]
    \begin{column}{0.4\textwidth}
      \begin{tabular}{|c|c|c|}
        \hline
        \textbf{Bid Qty}        & \textbf{Price}          & \textbf{Ask Qty}       \\
        \hline
        \cellcolor{gray!10}     & \cellcolor{gray!10}130  & \cellcolor{gray!10}24  \\
        \cellcolor{gray!10}     & \cellcolor{gray!10}125  & \cellcolor{gray!10}20  \\
        \cellcolor{gray!10}     & \cellcolor{gray!10}120  & \cellcolor{gray!10}10  \\
        \cellcolor{gray!10}15   & \cellcolor{gray!10}80   & \cellcolor{gray!10}    \\
        \cellcolor{gray!10}15   & \cellcolor{gray!10}75   & \cellcolor{gray!10}    \\
        \cellcolor{gray!10}30   & \cellcolor{gray!10}70   & \cellcolor{gray!10}    \\
        \hline
      \end{tabular}
    \end{column}

    \begin{column}{0.6\textwidth}
      \textbf{Discussion Questions}
      \begin{itemize}
        \item Why are all bids prices below all ask prices?
        \item Why do the quantities of quotes decrease as we approach the middle price?
        \item Would we consider this orderbook liquid?
        \item What is the market price of AMZN in this orderbook?
      \end{itemize}
    \end{column}
  \end{columns}

\end{frame}

\section{Trading Industry}

\begin{frame}{The Exchange Business Model}
  Exchanges make money by charging a fee per trade.

  Exchanges benefit from high trading volume, so they have an incentive to attract participants to their platform.

  \begin{sqt:result}
    To attract customers, exchanges:
    \begin{itemize}
      \item provide a fair and transparent trading environment.
      \item maintain high liquidity (by attracting market makers).
    \end{itemize}
  \end{sqt:result}
\end{frame}

\begin{frame}{Market Makers}
  \begin{sqt:definition}
    A \textbf{market maker} is a participant in an exchange who provides liquidity by continuously quoting buy and sell prices for a security.
  \end{sqt:definition}

  Market makers must fulfill obligations to the exchange, and maintain a good relationship with them.

  \begin{itemize}
    \item Quote a minimum size at all times.
    \item Quote within a certain spread.
    \item Don't do disorderly or manipulative trading.
  \end{itemize}

  In return, the exchange rewards market makers with:

  \begin{itemize}
    \item fee rebates
    \item not getting banned
  \end{itemize}
\end{frame}

\begin{frame}{Automated Trading}
  Programs run on machines that receive information from the exchange.
  The information allows them to build a view of the orderbook.
  They send messages to the exchange

  TODO: diagram

\end{frame}

\begin{frame}{High Frequency Trading}
  Any exchange activity that reacts to orderbook activity quicky.
  Medium Frequency is order of magnitude from milliseconds to seconds.

  TODO: talk about queue position
  TODO: talk about adverse selection
  TODO: finish

  Then talk about colocation.
\end{frame}

\begin{frame}{Buy Side and Sell Side}
  \begin{sqt:definition}
    \begin{itemize}
      \item The \textbf{buy side} consists of traders who buy exchange services. \\
      Liquidity is one of the most important of these services.
      \item The \textbf{sell side} consists of entities that provide these services.
    \end{itemize}
  \end{sqt:definition}

  TODO: give an example
\end{frame}

\begin{frame}{The Buy Side}
  TODO: investors, borrowers, hedgers, gamblers
  \begin{itemize}
    \item stuff
  \end{itemize}
  NOTE: they probably do not understand what hedging would mean
\end{frame}

\begin{frame}{The Sell Side}
  TODO: brokers, dealers, market makers
\end{frame}

\begin{frame}{Hedge Funds}
  \begin{sqt:definition}
    A \textbf{hedge fund} holds investor money and assets.
    An \textbf{investment management firm} trades the assets in the fund and makes returns,
    taking fees.
  \end{sqt:definition}

  Examples of hedge funds:
  \begin{itemize}
    \item Citadel
    \item QRT
  \end{itemize}
\end{frame}

\section{Trading \& Research}

\begin{frame}{Trader Roles}
  \begin{sqt:definition}
    Traders make decisions TODO: finish
  \end{sqt:definition}

  In HFT, traders adjust the parameters of the autotrader.
  They also write a lot of random scripts.
\end{frame}

\begin{frame}{Trader Interviews}
  Mental math speed (OA)
  Probability \& combinatorics
  Estimation (``make me a market on X'')
  Sometimes trading games (jane street definitely does this)
  Python (Jane Street definitely)
\end{frame}

\begin{frame}{Researcher}
  There is much more to research than just alpha and pricing theory.

  There are a wide variety of research roles, but they generally fall into the following categories:

  \begin{itemize}
    \item \textbf{Alpha}: Discovering predictive signals and trading opportunities
    \item \textbf{Pricing}: Pricing exotics (e.g. barrier options)
    \item \textbf{Risk}: Risk
    \item \textbf{Execution}: Transaction costs analysis, microstructure research
    \item \textbf{Portfolio Optimization}:
  \end{itemize}

  Roles also look different depending on time-horizon.

  TODO: finish this slide later
\end{frame}

\begin{frame}{Researcher Interviews: Core Skills}
  These interviews also have probability questions.
  Also may have questions about statistics or take home projects.

  Linear regression, generalized linear models, time series
\end{frame}

\begin{frame}{Trading Interview: Question Example 1}
  TODO: just pick something someone sent in the discord
\end{frame}

\section{Engineering}

\begin{frame}{Engineering Roles}

  \begin{itemize}
    \item \textbf{Software Engineer}
    \item \textbf{Quantitative Developer}
    \item \textbf{Hardware Engineer}
  \end{itemize}
\end{frame}

\begin{frame}{Software Engineering Roles}
  \begin{sqt:result}
    In trading firms, software engineers do much of the same work as software engineers in other industries.
  \end{sqt:result}
  \begin{itemize}
    \item \textbf{Application Development}
    \item \textbf{UI Development}
    \item \textbf{Data Engineering}
  \end{itemize}
\end{frame}

\begin{frame}{Software Engineering Interviews}
  Skills assessed:
  \begin{itemize}
    \item \textbf{Data Structures \& Algorithms}: binary search, graphs, complexity analysis
    \item \textbf{Design Skills}: System design, API design, OOP
    \item \textbf{General Knowledge}: Databases, Networks, OS, Concurrency, Compilers
    \item \textbf{Testing}: Unit tests, mock objects, designing for testability
  \end{itemize}
\end{frame}

\begin{frame}{Software Engineering Specialized Tracks}
  Many firms separate candidates into specialized tracks early in the process.

  \vspace{0.3cm}
  \textbf{Low-Latency Track}
  \begin{itemize}
    \item C++
    \item Operating Systems
    \item Computer Architecture
  \end{itemize}

  \vspace{0.3cm}
  \textit{Example firms: Citadel Securities, QRT}
\end{frame}

\begin{frame}{Software Engineering Specialized Tracks}
  \textbf{Python Track}
  \begin{itemize}
    \item Probability and statistics
    \item Data manipulation (NumPy, Pandas, SQL)
    \item Other stats (machine learning, data science)
  \end{itemize}

  \textit{Example firms: Citadel Securities}
\end{frame}

\begin{frame}{Quantitative Development}
  Half developer, half researcher.

  \vspace{0.3cm}
  \textbf{Math/Stats Skills}
  \begin{itemize}
    \item Undergrad math (linear algebra, probability, statistics)
    \item ML techniques (linear regression, clustering, decision trees, gradient boosting)
    \item Python for research (NumPy, Pandas, sklearn, statsmodels, plotting libraries)
  \end{itemize}

  \vspace{0.3cm}
  \textbf{Dev Skills}
  \begin{itemize}
    \item Expected to be as proficient as any other dev
  \end{itemize}

  \vspace{0.3cm}
  \textit{Example firms: Citadel Securities}
\end{frame}

\section{Interview Questions}

\begin{frame}{Probability Question 1}
  stuff...
\end{frame}

\section{Resources}

\begin{frame}{Resources}
  \textbf{Trading Fundamentals Talks}
  \begin{itemize}
    \item \href{https://www.youtube.com/watch?v=hYmR5qk4z8w}{Oden Petersen \& QuantSoc - Systematic Trading Workshop}
    \item \href{https://www.youtube.com/watch?v=2YkK4DnEBRE}{George Mukun Wang \& SQT - Intro to Quant Trading}
  \end{itemize}

  \textbf{Trading Interview Prep}
  \begin{itemize}
    \item Green book
    \item Dice problems
  \end{itemize}
\end{frame}

\begin{frame}{Resources}
  \textbf{Math} (I might just be a Botev glazer)
  \begin{itemize}
    \item \textit{An Advanced Course in Probability and Stochastic Processes - Zravko Botev}
    \item \textit{Data Science and Machine Learning - Zdravko Botev}
  \end{itemize}
\end{frame}

\begin{frame}{Resources}
  \textbf{Software Engineering Classics}
  \begin{itemize}
    \item \textit{Designing Data-Intensive Applications} (DDIA)
    \item \textit{Operating Systems: Three Easy Pieces} (OSTEP)
  \end{itemize}

  \textbf{C++ Classics}
  \begin{itemize}
    \item \textit{Exceptional C++} by Herb Sutter
    \item \textit{Effective Modern C++} by Scott Meyers
  \end{itemize}

  \textbf{Blogs, Websites, Videos}
  \begin{itemize}
    \item \href{https://agner.org}{Agner Fog Website}
  \end{itemize}
\end{frame}

\end{document}